---
title: "How to explore correlations?"
output:
  html_document:
    df_print: paged
---

[Source](https://drsimonj.svbtle.com/exploring-correlations-in-r-with-corrr)
```{r}
d <- mtcars
d$hp[3] <- NA
head(d)
```

We could be motivated by multicollinearity:

```{r}
fit_1 <- lm(mpg ~ hp,        data = d)
fit_2 <- lm(mpg ~ hp + disp, data = d)
```

```{r}
summary(fit_1)
```

```{r}
summary(fit_2)
```

Strange result. Let’s check the correlations between `mpg`, `hp`, and `disp` to try and diagnose this problem. It should be simple using the base R function, `cor()`. Right?

Err, what is with all the `NA`‘s ?

```{r}
rs <- cor(d)
rs
```

Have to handle missing values with `use`:

```{r}
rs <- cor(d, use = "pairwise.complete.obs")
rs
```

Can we focus on subset with dplyr? Nope.
```{r}
#dplyr::select(rs, mpg, hp, disp)
```
Riiiiiight! It’s a matrix and dplyr is for data frames.

```{r}
class(rs)
```

So we can use square brackets with matrices? Or not…

```{r}
vars <- c("mpg", "hp", "disp")
rs[rownames(rs) %in% vars]
```
Mm, square brackets can take on different functions with matrices. Without a comma, it’s treated like a vector. With a comma, we can separately specify the dimensions.

```{r}
vars <- c("mpg", "hp", "disp")
rs[rownames(rs) %in% vars, colnames(rs) %in% vars]
```

We diagnosed our multicollinearity problem. What if we want to something a bit more complex like exploring clustering of variables in high dimensional space? Could use exploratory factor analysis.

```{r}
factanal(na.omit(d), factors = 2)
```

```{r}
factanal(na.omit(d), factors = 5)
```

So many questions! I’d much rather explore the correlations.

Let’s try to find all variables with a correlation greater than 0.90. Why doesn’t this work?!

```{r}
col_has_over_90 <- apply(rs, 2, function(x) any(x > .9))
rs[, col_has_over_90]
```

The diagonal is 1. All cols have a value greater than .90!

Exclude diagonal:

```{r}
diag(rs) <- NA
col_has_over_90 <- apply(rs, 2, function(x) any(x > .9, na.rm = TRUE))
rs[, col_has_over_90]
```

## Exploring data with the tidyverse

```{r}
library(tidyverse)
d %>% 
  select(mpg:drat) %>% 
  gather() %>% 
  ggplot(aes(value)) +
    geom_histogram() +
    facet_wrap(~key, scales = "free")
```

# Using `corr`

```{r}
library(corrr)
d %>% 
  correlate() %>% 
  focus(mpg:drat, mirror = TRUE) %>% 
  network_plot()
```

```{r}
rs <- correlate(d)
rs
```

```{r}
rs %>% 
  select(mpg:drat) %>% 
  gather() %>% 
  ggplot(aes(value)) +
    geom_histogram() +
    facet_wrap(~key)
```

How about that challenge to find cols with a correlation greater than .9?
```{r}
any_over_90 <- function(x) any(x > .9, na.rm = TRUE)
rs %>% select_if(any_over_90)
```

```{r}
rs %>% 
  focus(mpg, disp, hp)
```

```{r}
rs %>% 
  focus(-mpg, -disp, -hp)
```

```{r}
rs %>% 
  focus(mpg, disp, hp, mirror = TRUE)
```

```{r}
rs %>% 
  focus(matches("^d"))
```

```{r}
rs %>% 
  focus(mpg)
```

```{r}
rs %>%
  focus(mpg) %>%
  mutate(rowname = reorder(rowname, mpg)) %>%
  ggplot(aes(rowname, mpg)) +
    geom_col() + coord_flip()
```

```{r}
rs %>% rearrange()
```

```{r}
rs %>% shave()
```

```{r}
rs %>% stretch()
```

```{r}
rs %>%
  shave() %>% 
  stretch(na.rm = FALSE) %>% 
  ggplot(aes(r)) +
    geom_histogram()
```

```{r}
rs %>%
  focus(mpg:drat, mirror = TRUE) %>% 
  rearrange() %>% 
  shave(upper = FALSE) %>% 
  select(-hp) %>% 
  filter(rowname != "drat")
```

```{r}
rs %>% fashion()
```

```{r}
rs %>%
  focus(mpg:drat, mirror = TRUE) %>% 
  rearrange() %>% 
  shave(upper = FALSE) %>% 
  select(-hp) %>% 
  filter(rowname != "drat") %>% 
  fashion()
```

```{r}
rs %>% rplot()
```

```{r}
rs %>%
  rearrange(method = "MDS", absolute = FALSE) %>%
  shave() %>% 
  rplot(shape = 15, colors = c("red", "green"))
```

```{r}
rs %>% network_plot(min_cor = .6)

```

